---
title: "`r tools::toTitleCase(gsub('_',' ',tools::file_path_sans_ext(knitr::current_input())))`"
author: Rui Xiao
date: "`r Sys.Date()`"
params:
  f: "cl0814.tsv"
  x: "1"
  y: "2:ncol(df)"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      hidden: true
    number_sections: true
    # code_folding: hide
    theme: readable
    highlight: haddock
    self_contained: true
    css: styles.css
  pdf_document:
    latex_engine: xelatex
    toc: true
    toc_depth: 2
    number_sections: true
# ```{r, child="_base.rmd", eval=T}
# ```
# word_document
---


# Initialize Environment
```{r option, include=FALSE}
knitr::opts_chunk$set(echo=F,warning=F,message=F,tidy=T,
dpi=150,fig.width=5,fig.height=5,out.width="720px",out.height="720px")
is_html<-knitr::is_html_output(exclude=c("markdown","epub"))
```

```{r param, out.height="12"}
suppressMessages(library(tidyverse))
library(readxl)
if (is_html) {
  library(DT)
  library(plotly); plot_ly()
}
source("_helper.R")
print(params)
```

# Data Processing
```{r read}
pf<-params$f
if (is.null(pf) || !file.exists(pf)) stop("no input file or file doesn't exist")
# df<-read_tb(pf,na=c("NaN","NA",""))
df<-read_tb(pf,na=c("NaN","NA","",0)) # WARNING, also exclude ZERO
if (is_html) {
  DT::datatable(df, escape=F, rownames=F, filter="top", fillContainer=T,
  options=list(pageLength=5, scrollX=T, scrollY=F))
} else {
  cat("Actual dimension: [",paste(dim(df),collapse="x"),"], preview:\n",sep=""); print(head(df))
}

px<-ifelse(is.null(params$x), params$x, "1")
py<-ifelse(is.null(params$y), params$y, "2:ncol(df)")

(xs<-colnames(df)[eval(parse(text=px))])
x<-xs[1]; gs<-sort(unique(df[[x]])) # only 1 y, split into groups
(ys<-colnames(df)[eval(parse(text=py))])

```

# Select Groups
```{r selgroup, results='asis'}
tab<-freq_pct(df[[x]]) %>% rownames_to_column(var="Group") # only 1st y con
cat("\n```\n", sep=""); print(tab); cat("\n```\n", sep="")
tab$Group<-fac_wrap_rev(tab$Group)
if (is_html) {
  p<-plot_ly(tab, type='bar', orientation='h',
      x=~Freq, y=~Group, color=~Group,
      text=~Freq, textposition='outside', textfont=list(color='black')) %>%
    layout(xaxis=list(title="cases"), yaxis=list(title=""),
      legend=list(traceorder="reversed"))
  print(htmltools::tagList(p))
} else {
  p<-ggplot(tab, aes(x=Group, y=Freq, fill=Group, label=Freq))+
    geom_col(width=0.6,alpha=0.7)+geom_text()+
    guides(fill=guide_legend(reverse=T))+
    theme_bw()+labs(x="")+coord_flip()
  print(p)
}
```

# Each Comparison
```{r loopCol, results='asis', warning=F}
for (y in ys){
  cont<-is_cont(df[[y]])
  cat("\n\n## ",y, "\n",
    "\n```\n", y, " (", ifelse(cont, "continuous", "categorical"), ")",
    sep="")
  dsv<-df[is_valid(df[[y]]), c(x,y)]; colnames(dsv)<-c("Group","Var")
  if (cont) { # continuous
    res<-dplyr::tibble(Group=character(), n=numeric(),
      y=numeric(), ymin=numeric(), ymax=numeric(), Label=character())
    for (g in gs) {
      dat<-dsv$Var[dsv$Group==g]; dat<-dat[is_valid(dat)] # group specific valid datapoints
      d_n<-length(dat); d_m<-mean(dat); d_sd<-sd(dat)
      res<-dplyr::add_row(res, Group=g, n=d_n, y=d_m, ymin=d_m-d_sd, ymax=d_m+d_sd,
        Label=paste0(round(d_m,1), "±", round(d_sd,1), " (", d_n, ")"))
    }
    print(quantile(dsv$Var)); print(res)
    form<-as.formula(paste("Var","Group",sep="~"))
    if (length(gs)==2) { # 2 groups
      cat("t-test (Welch Two Sample): P ", ps_fmt(t.test(form, data=dsv)$`p.value`), "\n", sep="")
    } else { # 3+ groups
      cat("ANOVA summary:\n"); aov<-aov(form, data=dsv); print(summary(aov))
      cat("Tukey multiple comparisons (95% CI):\n")
      tukey<-as.data.frame(TukeyHSD(x=aov)$Group)
      tukey$signif<-ps_sym(tukey$`p adj`)
      print(tukey)
    }
    cat("\n```\n")
    res$Group<-fac_wrap_rev(paste(y,res$Group))
    dsv$Group<-fac_wrap_rev(paste(y,dsv$Group))
    if (is_html) { # continuous html
      p<-plot_ly(dsv, x=~Var, y=~Group, split=~Group, type='violin', orientation='h',
        box=list(visible=T), meanline=list(visible=T), points="outliers",
        opacity=0.75, height=20+120*nlevels(res$Group)) %>%
      add_annotations(xref="Var", yref="Group", x=res$y, y=res$Group,
        text=gsub(" ","\n",res$Label), ax=0) %>%
      layout(title=list(text=y), xaxis=list(title="mean, sd, (sample size)"), yaxis=list(title=""),
        legend=list(traceorder="reversed"))
      print(htmltools::tagList(p))
    } else { # continuous flat
      p<-ggplot(dsv, aes(x=Group, y=Var, color=Group, fill=Group))+
        geom_violin(trim=F, alpha=0.2, width=0.4)+
        geom_pointrange(aes(y=y, ymin=ymin, ymax=ymax), data=res, show.legend=F)+
        geom_text(aes(y=y, label=gsub(" ","\n",Label)), data=res, color="black", show.legend=F)+
        labs(x="",y="mean, sd, (sample size)", title=y)+ theme_bw()+coord_flip()
      print(p)
    }
  } else { # categorical
    res<-dplyr::tibble(Group=character(), Category=character(), n=numeric(),
      y=numeric(), ymin=numeric(), ymax=numeric(), Label=character())
    for (g in gs) {
      for (v in sort(unique(dsv$Var))) {
        match_ctg<-dsv$Var[dsv$Group==g]==v
        br<-boot::boot(match_ctg, pct_true, R=100)
        brp<-boot::boot.ci(br, type='perc')$percent
        res<-dplyr::add_row(res, Group=g, Category=as.character(v), n=sum(match_ctg),
        y=br$t0, ymin=brp[4], ymax=brp[5], Label=paste0(round(br$t0*100,1),"% (",sum(match_ctg),")"))
      }
    }
    if (nrow(res)>0) {
      print(res); print(chisq_multi(res))
      cat("\n```\n")
      if (length(unique(res$Category))>1) {
        res$Group<-fac_wrap_rev(paste(y,res$Group))
        res$Category<-fac_wrap_rev(res$Category)
        if (is_html) { # categorical html
          res<-res[with(res, order(Category,Group)), ] # align error bar
          p<-plot_ly(res, x=~y, y=~Group, color=~Category, text=~gsub(" ","\n",Label),
            height=10+60*(nlevels(res$Group)+nlevels(res$Category)),
            textposition='outside', textfont=list(color="black"),
            error_x=~list(symmetric=F, y=~Group, array=ymax-y, arrayminus=y-ymin, thickness=3),
            marker=list(opacity=0.55, line=list(color="black",width=2)), type='bar', orientation='h') %>%
          layout(xaxis=list(title="%, 95% CI, (sample size)"), yaxis=list(title=""), title=list(text=y),
            legend=list(traceorder='reversed'), barmode='group')
          print(htmltools::tagList(p))
        } else { # categorical flat
          p<-ggplot(res, aes(x=Group, y=y, group=Category, color=Category,shape=Category))+
            geom_pointrange(aes(ymin=ymin, ymax=ymax), position=position_dodge(0.9))+
            geom_line(position=position_dodge(0.9))+
            geom_text(aes(label=Label), color="black", position=position_dodge(0.9), show.legend=F)+
            labs(x="",y="%, 95% CI, (sample size)", title=y)+
            guides(color=guide_legend(reverse=T), shape=guide_legend(reverse=T))+
            scale_y_continuous(labels=scales::percent, expand=expansion(mult=c(0.1,0.1)))+
            theme_bw()+coord_flip()
          print(p)
        }
      } else { cat("\n```\n")}
    }
  }
  #break
}

```

